---
title: "Question 5"
output: html_notebook
---
##(5). Finally, you must also illustrate the prediction ability of the developed model using testing set. Think of creative ways to convey this to your client (Rohan).

```{r}
model_final <- lm(Price ~ Kilometers_Driven + Mileage + Power + Seats + Fuel_Type 
+ Transmission, data = traincars)

predictions <- predict(model_final, newdata = testcars)
  
library(ggplot2)
ggplot(data = testcars, aes(x = Price, y = predictions)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, intercept = 0, color = "magenta", linetype = "dashed") +
  labs(title = "Actual vs Predicted Prices",
       x = "Actual Prices",
       y = "Predicted Prices") 

```

```{r}
predictions <- predict(model_final, newdata = testcars)
actual <- testcars$Price
mae <- mean(abs(actual - predictions))
rmse <- sqrt(mean((actual - predictions)^2))

print(mae)
print(rmse)
```

Using our model, we created a graph to show Rohan that our model accuratly predicts the price of the used cars. Our model demonstrates strong predictive ability, as validated through testing set results. It explains **81.53% of the variability in vehicle prices** (adjusted \(R^2\)) and has consistent residual patterns that confirm linear regression assumptions. Key predictors like **Kilometers Driven**, **Mileage**, **Power**, **Fuel Type**, and **Transmission** are statistically significant, ensuring the model captures meaningful relationships. While errors (MAE = ₹312,135 and RMSE = ₹399,095) indicate some variability, these values remain reasonable given the range of vehicle prices. Visualization of actual vs. predicted prices further illustrates the model’s reliability. The model is interpretable and user-friendly, making it a valuable tool for Rohan's pricing decisions.
